STR: Seasonal-Trend decomposition using Regression

### 2021-04-04 str_submission.tex
Updated version for submission

\graphicspath{{figs/}}
\def\addlinespace{}
\usepackage{amsmath,bm,hyperref,url,booktabs}
\setlength{\arraycolsep}{3pt}
\setcounter{MaxMatrixCols}{20}
\renewcommand{\vec}{\operatorname{vec}}
\newcommand{\diag}{\operatorname{diag}}
\renewcommand{\mod}{\operatorname{mod}}
\usepackage[section]{placeins}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

@@

% Block of authors and their affiliations starts here:
% NOTE: Authors with same affiliation, if the order of authors allows,
%   should be entered in ONE field, separated by a comma.
%   \EMAIL field can be repeated if more than one author
\ARTICLEAUTHORS{%
\AUTHOR{Alexander Dokumentov}
\AFF{Growing Data, Australia} %, \URL{}}
\AUTHOR{Rob J Hyndman}
\AFF{Department of Econometrics \& Business Statistics, Monash University, Clayton VIC 3800, Australia, \EMAIL{Rob.Hyndman@monash.edu}}
% Enter all authors
} % end of the block

@@

% Fill in data. If unknown, outcomment the field
\KEYWORDS{time series decomposition, seasonal data, Tikhonov regularisation, ridge regression, LASSO, STL, TBATS, X-12-ARIMA, BSM}

@@

\section{Time series decomposition}\label{time-series-decomposition}

@@

The first attempts to decompose time series into various components can be dated to the 19th century, when \citet{Poynting1884} proposed price averaging as a tool for eliminating trend and seasonal fluctuations. Later, his approach was extended by \citet{hooker1901suspension}, \citet{spencer1904graduation} and \citet{anderson1914nochmals}. Further research in this direction included that by \citet{copeland1915statistical}, who was the first to attempt to extract the seasonal component, until \citet{macaulay1930smooth} proposed a method which gradually became ``classical''. The work of Macaulay led to the Census II method \citep{shiskin1957electronic} which became popular amongst national statistics offices. The Census II method has continued to evolve, with additional features including robustness, calendar effects, covariates, ARIMA extensions, and extensive diagnostics. Widely used versions of this approach have been X-11 \citep{shishkin1967x}, X-11-ARIMA \citep{dagum1988x11arima,ladiray2001seasonal}, X-12-ARIMA \citep{findley1998new} and X-13-ARIMA-SEATS \citep{findley2005some}.

@@

Several model-based methods for seasonal decomposition have been developed, including the TRAMO/SEATS procedure which was developed at the Bank of Spain \citep{Gomez2001}, the TBATS model of \citet{de2011forecasting}, and various structural time series model approaches \citep{harvey1990forecasting,Commandeur2011}. One big advantage of using a model for seasonal decomposition and adjustment is that it provides a natural way to compute confidence and prediction intervals.

@@

@@

@@

@@

\begin{figure}[!b]\FIGURE
{\includegraphics[width=.9\textwidth]{str_ijds_files/figure-latex/supermarket-1.pdf}}
{\label{fig:supermarket}Supermarket and grocery store log revenue in New South Wales, decomposed with STR. The original data (logged) are shown in the top panel, while the components are shown in the other panels, with 95\% confidence intervals for the trend and season components in blue}
{}
\end{figure}

@@

We obtained similar results using other decomposition methods applied to the same data including STL \citep{cleveland1990stl} and X-13-ARIMA-SEATS \citep{findley2005some}.

%\FloatBarrier

@@

\begin{figure}\FIGURE
{\includegraphics[width=.9\textwidth]{str_ijds_files/figure-latex/vic-1.pdf}}
{\label{fig:vic}Electricity demand in Victoria, decomposed using STR. The original data are plotted in black in the first panel, with the decomposed trend plotted in red. The subsequent panels contain the weekly seasonal pattern; the daily seasonal pattern with complex topology; the effect of temperature in Melbourne, which has a daily seasonal pattern with complex topology; the effect of squared temperatures in Melbourne, which is allowed to vary over time; and the residuals. Grey vertical lines represent the residuals with the largest absolute values.}
{}
\end{figure}

@@

\begin{table}\TABLE
{\label{tab:outlierstable}The ten residuals that are largest in absolute value after an STR decomposition.}
{\begin{tabular}[t]{rlcr}
\toprule
Date & Day of week & Time period & Residual\\
\midrule
3 February 2000 & Thursday & 14:30 -- 15:00 & 687.3\\
3 February 2000 & Thursday & 15:00 -- 15:30 & 871.3\\
3 February 2000 & Thursday & 15:30 -- 16:00 & 937.0\\
3 February 2000 & Thursday & 16:00 -- 16:30 & 767.2\\
3 February 2000 & Thursday & 16:30 -- 17:00 & 580.4\\

4 February 2000 & Friday & 16:00 -- 16:30 & -569.7\\
10 February 2000 & Thursday & 14:30 -- 15:00 & -611.4\\
10 February 2000 & Thursday & 15:00 -- 15:30 & -658.9\\
10 February 2000 & Thursday & 15:30 -- 16:00 & -695.7\\
10 February 2000 & Thursday & 16:00 -- 16:30 & -728.5\\
\bottomrule
\end{tabular}}
{}
\end{table}

\begin{figure}\FIGURE
{\includegraphics[width=\textwidth]{str_ijds_files/figure-latex/MelTemp-1.pdf}}
{{\label{fig:MelTemp}The temperature in Melbourne on 1 February 2000, and the following ten days. Ten red lines (two groups of five lines each) mark the times of the outliers. The first group spans 14:30--17:00 on 3 February, while the second group spans 15:30--18:00 on the following day.}}
{}
\end{figure}

@@

Similarly, seasonal functional components can be obtained, using Fourier terms for example \citep[reminiscent of][]{de2011forecasting}. This approach leads, with no additional effort, to handling seasonality with a fractional or varying period.

@@

>>>>

\graphicspath{{figs/}}
\def\addlinespace{}
\usepackage{amsmath,bm,hyperref,url,booktabs}
\usepackage{pdfpages}
\setlength{\arraycolsep}{3pt}
\setcounter{MaxMatrixCols}{20}
\renewcommand{\vec}{\operatorname{vec}}
\newcommand{\diag}{\operatorname{diag}}
\renewcommand{\mod}{\operatorname{mod}}
\usepackage[section]{placeins}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

@@

% Block of authors and their affiliations starts here:
% NOTE: Authors with same affiliation, if the order of authors allows,
%   should be entered in ONE field, separated by a comma.
%   \EMAIL field can be repeated if more than one author
\ARTICLEAUTHORS{%
\AUTHOR{Alexander Dokumentov}
\AFF{Let's Forecast, Australia} %, \URL{}}
\AUTHOR{Rob J Hyndman}
\AFF{Department of Econometrics \& Business Statistics, Monash University, Clayton VIC 3800, Australia, \EMAIL{Rob.Hyndman@monash.edu}}
% Enter all authors
} % end of the block

@@

% Fill in data. If unknown, outcomment the field
\KEYWORDS{time series decomposition, seasonal data, Tikhonov regularisation, ridge regression, LASSO, STL, TBATS, X-12-ARIMA, X-13-ARIMA-SEATS, BSM}

@@

\hypertarget{time-series-decomposition}{%
\section{Time series decomposition}\label{time-series-decomposition}}

@@

The first attempts to decompose time series into various components can be dated to the 19th century, when \citet{Poynting1884} proposed price averaging as a tool for eliminating trend and seasonal fluctuations. Later, his approach was extended by \citet{hooker1901suspension}, \citet{spencer1904graduation} and \citet{anderson1914nochmals}. Further research in this direction included that by \citet{copeland1915statistical}, who was the first to attempt to extract the seasonal component, until \citet{macaulay1930smooth} proposed a method which gradually became ``classical''. The work of Macaulay led to the Census II method \citep{shiskin1957electronic} which became popular amongst national statistics offices. The Census II method has continued to evolve, with additional features including robustness, calendar effects, covariates, ARIMA extensions, and extensive diagnostics. Widely used versions of this approach have been X-11 \citep{shishkin1967x}, X-11-ARIMA \citep{dagum1988x11arima,ladiray2001seasonal}, X-12-ARIMA \citep{findley1998new} and X-13-ARIMA-SEATS \citep{findley2005some,Dagum2016}.

@@

Several model-based methods for seasonal decomposition have been developed, including the TRAMO/SEATS procedure which was developed at the Bank of Spain \citep{Gomez2001}, the TBATS model of \citet{delivera2011forecasting}, and various structural time series model approaches \citep{harvey1990forecasting,Commandeur2011}. One big advantage of using a model for seasonal decomposition and adjustment is that it provides a natural way to compute confidence and prediction intervals.

@@

Note that the model subsumes the case of a linear trend, obtained when \(\sigma_L=0\) (making the trend's second derivatives equal to zero).

@@

Again, by setting the smoothing parameters to zero, some simple special cases are obtained. Setting the variance of the seasonal component in the time-seasonal direction to zero causes the seasonal component to be periodic. By setting both trend and time-seasonal variances to zero we effectively force our model to have a linear trend with seasonal dummies. Setting the variance of the second derivatives of the seasonal component in the time direction to zero alone makes changes in the seasonal component linear over time, but does not preclude the seasonal pattern from changing.

@@

\hypertarget{simulations}{%
\section{Simulations}\label{simulations}}

To demonstrate that our method estimates the appropriate time series components, we use simulations where we know the true underlying components. We will compare our results against two other decomposition methods that handle multiple seasonal components, STL and TBATS. STL \citep{cleveland1990stl} is a well known decomposition method, and therefore a natural choice to use as a method to compare. The original methodology and implementation was designed for only one seasonal period. However, the implementation in the forecast package for R \citep{Rforecast} will compute an STL decomposition iteratively with multiple seasonal periods. TBATS \citep{delivera2011forecasting} is a forecasting method designed for multiple seasonalities; it provides a decomposition of the time series into components, and so can provide an alternative comparison.

Our simulated daily data consists of four components:
\begin{equation}
  \label{STR:many_parts2}
  y_{t} = T_{t} + \alpha S^{W}_{t} + \beta S^{Y}_{t} + \gamma R_{t} , \quad t=1,\dots,n,
\end{equation}
where \(T_{t}\) is a trend, \(S^{W}_{t}\) is a weekly seasonal component, \(S^{Y}_{t}\) is a yearly seasonal component, \(R_{t}\) is the remainder, and \(\alpha\), \(\beta\), and \(\gamma\) are parameters which control the contribution of the components to \(y_{t}\).

We use two data generating processes (DGPs) to simulate data. For the first DGP, we use a quadratic trend function with random coefficients, such that \(T_t^* = N_1(t + n/2(N_2-1))^2\) where \(N_1\) and \(N_2\) are independent N(0,1) random variables. Then \(T_t^*\) is normalized to give \(T_t\) with mean zero and unit variance. The seasonal components consist of five pairs of Fourier terms with random N(0,1) coefficients, which are then also normalized. Because this DGP has deterministic components, we refer to it as the ``Deterministic'' GDP.

For the second ``Stochastic'' DGP, the trend is given by an ARIMA(0,2,0) model with standard normal errors, which is then normalized. The seasonal components are also smooth and obtained using a similar process constrained to be periodic. Specifically, for each seasonal component we start with a vector of iid N(0,1) values of the length of one season, normalized and replicated to be of length \(n\). This is then integrated and normalized twice to make it smooth. So each seasonal component is as smooth as the trend, is centred on zero and is periodic.

Figure \ref{fig:examples} shows examples of simulated time series from each of the two DGPs with \(\alpha=\beta=1\) and \(\gamma=0.25\). We generated 1096 days in each case (one day more than three non-leap years of daily data).

\begin{figure}
\centering
\includegraphics[width=\textwidth]{str_ijds_files/figure-latex/examples-1.pdf}
\caption{\label{fig:examples}Examples of time series generated by the two DGPs.}
\end{figure}

For comparison of the three decomposition methods, we use parameter sets \((\alpha, \beta, \gamma) = \{(1, 1, 0.2), (1, 1, 0.4), (1, 1, 0.6)\}\), and create 20 time series for each DGP and parameter set. We apply STR, STL and TBATS to each series and compute RMSE for each component (averaging across all data sets and days). We measure the significance of the differences between the methods using a linear model applied to the squared errors. The results are shown in Table \ref{tab:cbetable}. It is clear that STR significantly outperforms STL and TBATS in estimating all components.

\begin{table}[!ht]

\caption{\label{tab:cbetable}RMSE over 100 simulations for each DGP and value of $\gamma$. Bold values indicate results that are significantly different from the STR values.}
\centering
\begin{tabular}[t]{rlrrrr}
\toprule
$\gamma$ & Method & Trend RMSE & Weekly RMSE & Yearly RMSE & Remainder RMSE\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Stochastic DGP}}\\
\hspace{1em}0.2 & STL & \textbf{0.0808} & \textbf{0.0564} & \textbf{0.1267} & \textbf{0.1525}\\
\hspace{1em}0.2 & TBATS & \textbf{0.1830} & \textbf{0.0174} & \textbf{0.1767} & \textbf{0.0566}\\
\hspace{1em}0.2 & STR & 0.0183 & 0.0149 & 0.0475 & 0.0515\\
\midrule
\hspace{1em}0.4 & STL & \textbf{0.0798} & \textbf{0.1148} & \textbf{0.2299} & \textbf{0.2659}\\
\hspace{1em}0.4 & TBATS & \textbf{0.3435} & \textbf{0.0344} & \textbf{0.3310} & \textbf{0.1227}\\
\hspace{1em}0.4 & STR & 0.0320 & 0.0301 & 0.0778 & 0.0876\\
\midrule
\hspace{1em}0.6 & STL & \textbf{0.0846} & \textbf{0.1717} & \textbf{0.3388} & \textbf{0.3884}\\
\hspace{1em}0.6 & TBATS & \textbf{0.5107} & \textbf{0.0508} & \textbf{0.4983} & \textbf{0.1648}\\
\hspace{1em}0.6 & STR & 0.0523 & 0.0438 & 0.1136 & 0.1273\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Deterministic GDP}}\\
\hspace{1em}0.2 & STL & \textbf{0.1272} & \textbf{0.0575} & \textbf{0.1402} & \textbf{0.1748}\\
\hspace{1em}0.2 & TBATS & \textbf{0.2304} & \textbf{0.0166} & \textbf{0.2259} & \textbf{0.0803}\\
\hspace{1em}0.2 & STR & 0.0851 & 0.0147 & 0.1000 & 0.0607\\
\midrule
\hspace{1em}0.4 & STL & \textbf{0.1135} & \textbf{0.1147} & \textbf{0.2374} & \textbf{0.2757}\\
\hspace{1em}0.4 & TBATS & \textbf{0.3393} & \textbf{0.0340} & \textbf{0.3294} & \textbf{0.1153}\\
\hspace{1em}0.4 & STR & 0.0584 & 0.0306 & 0.0937 & 0.0929\\
\midrule
\hspace{1em}0.6 & STL & \textbf{0.1439} & \textbf{0.1714} & \textbf{0.3427} & \textbf{0.4018}\\
\hspace{1em}0.6 & TBATS & \textbf{0.5444} & \textbf{0.0509} & \textbf{0.5298} & \textbf{0.1709}\\
\hspace{1em}0.6 & STR & 0.0757 & 0.0436 & 0.1691 & 0.1709\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{str_ijds_files/figure-latex/supermarket-1.pdf}
\caption{\label{fig:supermarket}Supermarket and grocery store log revenue in New South Wales, decomposed with STR. The original data (logged) are shown in the top panel, while the components are shown in the other panels, with 95\% confidence intervals for the trend and season components in blue}
\end{figure}

@@

We obtained similar results using other decomposition methods applied to the same data including STL \citep{cleveland1990stl}, TBATS \citep{delivera2011forecasting} and X-13-ARIMA-SEATS \citep{findley2005some,Dagum2016}. These comparisons are provided in the appendix.

\FloatBarrier

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{str_ijds_files/figure-latex/vic-1.pdf}
\caption{\label{fig:vic}Electricity demand in Victoria, decomposed using STR. The original data are plotted in black in the first panel, with the decomposed trend plotted in red. The subsequent panels contain the weekly seasonal pattern; the daily seasonal pattern with complex topology; the effect of temperature in Melbourne, which has a daily seasonal pattern with complex topology; the effect of squared temperatures in Melbourne, which is allowed to vary over time; and the residuals. Grey vertical lines represent the residuals with the largest absolute values.}
\end{figure}

@@

\begin{table}[!bh]

\caption{\label{tab:outlierstable}The ten residuals that are largest in absolute value after an STR decomposition.}
\centering
\begin{tabular}[t]{rlcr}
\toprule
Date & Day of week & Time period & Residual\\
\midrule
3 February 2000 & Thursday & 14:30 -- 15:00 & 687.3\\
10
3 February 2000 & Thursday & 15:00 -- 15:30 & 871.3\\
10
3 February 2000 & Thursday & 15:30 -- 16:00 & 937.0\\
10
3 February 2000 & Thursday & 16:00 -- 16:30 & 767.2\\
10
3 February 2000 & Thursday & 16:30 -- 17:00 & 580.4\\
10
4 February 2000 & Friday & 16:00 -- 16:30 & -569.7\\
10
10 February 2000 & Thursday & 14:30 -- 15:00 & -611.4\\
10
10 February 2000 & Thursday & 15:00 -- 15:30 & -658.9\\
10
10 February 2000 & Thursday & 15:30 -- 16:00 & -695.7\\
10
10 February 2000 & Thursday & 16:00 -- 16:30 & -728.5\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{str_ijds_files/figure-latex/MelTemp-1.pdf}
\caption{\label{fig:MelTemp}The temperature in Melbourne on 1 February 2000, and the following ten days. Ten red lines (two groups of five lines each) mark the times of the outliers. The first group spans 14:30--17:00 on 3 February, while the second group spans 15:30--18:00 on the following day.}
\end{figure}

@@

Similarly, seasonal functional components can be obtained, using Fourier terms for example \citep[reminiscent of][]{delivera2011forecasting}. This approach leads, with no additional effort, to handling seasonality with a fractional or varying period.

@@

\hypertarget{feature-engineering}{%
\subsection{Feature engineering}\label{feature-engineering}}

The STR model provides a way to identify new features in time series data, such as weekly or daily patterns. These can then be used within other algorithms. For example, they could be used to classify time series \citep{catch22}, to find anomalous time series (as distinct from anomalous observations) \citep{cikm2015}, to select a forecasting model \citep{fforms}, or to create a weighted forecast combination \citep{fforma}.
